Supporting Information : Understanding the nanoscale structure of hexagonal phase lyotropic liquid crystal membranes

### 2018-09-15 Ben_Manuscripts/structure_paper/supporting_information.tex
revisions

\externaldocument[M-]{draft}
\title{Supporting Information : Understanding the nanoscale structure of hexagonal phase lyotropic liquid crystal membranes}
\author{Benjamin J. Coscia \and Douglas L. Gin \and Richard D. Noble \and Joe Yelk \and Matthew Glaser \and Xunda Feng \and Michael R. Shirts}

@@

  We parameterized monomers according to the following procedure:
  \begin{enumerate}
	\item \textit{Create monomer structure file with connectivity} : We
	drew atomistic structures using MarvinSketch
	17.13~\cite{chemaxon_marvinsketch_2017} with all hydrogen atoms drawn out
	explicitly. We optimized the 3D geometry of the structure using the 'Clean in
	3D' function of MarvinSketch.  We saved the structure as a .mol file, then
	converted it to .pdb format using Open Babel 2.4.1
	\cite{oboyle_open_2011,noauthor_open_nodate}. 
	\item \textit{Assign GAFF atomtypes using \texttt{antechamber}} : Using
	the .pdb structure file as input, we ran
	\texttt{antechamber}~\cite{wang_automatic_2006} using the AM1-BCC charge model.
	The net charge on the monomer is input as -1 since the sodium ion is kept as 
	a separate residue. We use \texttt{LEaP}~\cite{case_ambertools16_2016} and the
	output of \texttt{antechamber} to create Amber topology files. A detailed
	tutorial can be accessed elsewhere \cite{walker_antechamber_nodate}.
	\item \textit{Create GROMACS topologies from Amber output} : The output
	of \texttt{LEaP} is a .inpcrd and a .prmtop file which are Amber topology
	files. Using acpype.py \cite{sousa_da_silva_acpype_2012}, we converted the
	\texttt{LEaP} output into GROMACS .gro and .top files. 
	\item \textit{Perform a simulated annealing procedure on the monomer} :
	We created a cubic box around the monomer using the GROMACS command \texttt{gmx
	editconf}. The monomer was centered in the box with edges of the
	box spaced at least 3 nm from the monomer on all sides. We ran an energy minimization
	on the system with the steepest descent algorithm. Next we performed an NVT
	simulated annealing procedure. We linearly decreased the temperature of the
	system from 1000 K to 50 K over the course of 10 ns. We randomly chose a monomer
	configuration from the last 10 \% of the trajectory. 
	\item \textit{Reassign charges with \texttt{molcharge}}: With the monomer
	configuration taken from the annealed trajectory, we reassigned charges using
	\texttt{molcharge} with the am1bccsym method in order to ensure charges 
	are symmetric. This condition is not guaranteed with \texttt{antechamber}. 
	The charges in the GROMACS topology file (.top) were replaced with the 
	new charges calculated by \texttt{molcharge}. 
	\item \textit{Anneal again to get final structure} : We repeated the
	same simulated annealing procedure using the monomer topology with
	\texttt{molcharge} charges. A random monomer configuration was pulled from the
	last 10 \% of the trajectory and was used to build all assemblies reported
	(Figure~\ref{fig:monomer}).
  \end{enumerate} 
  \clearpage
  \begin{figure}[!htb]
	\centering
        \includegraphics[width=0.9\textwidth]{monomer.png}
	\caption{Atomistic representation of the monomer Na-GA3C11. White atoms
		represent hydrogen, cyan atoms represent carbon, red atoms represent oxygen and
		the blue atom is sodium.}\label{fig:monomer}
  \end{figure}

@@

  We use the nematic order parameter as defined in section~\ref{method:nematic_order} to 
  observe any progress towards system ordering. In both cases, we see that the nematic
  order parameter stays close to zero for the duration of the simulation time, indicating
  an isotropic arrangement of head groups (Figure~\ref{fig:nematic_isotropic}). The values
  are slightly negative since there are only 400 monomers. 
  Equation~\ref{eqn:nematic_order_parameter} is bounded between -1 and 1, however, with a
  proper choice of nematic director, measured values will range between 0 and 1. With a 
  large number of samples, an isotropic system would have an order parameter of 0.

@@

@@

	% BJC: Could put vertical lines to indicate where position restraint values are reduced
  	\begin{figure}[!htb]
	\centering
	\begin{subfigure}{0.3\textwidth}
		\includegraphics[width=\textwidth]{p2p_39.png}\quad
		\vspace{-1.25em}
		\caption{39 \AA~initial pore spacing}~\label{fig:p2p_39}
	\end{subfigure}
	\begin{subfigure}{0.3\textwidth}
		\includegraphics[width=\textwidth]{p2p_41.png}\quad
		\vspace{-1.25em}
		\caption{41 \AA~initial pore spacing}~\label{fig:p2p_41}
	\end{subfigure}
	\begin{subfigure}{0.3\textwidth}
		\includegraphics[width=\textwidth]{p2p_45.png}
		\vspace{-1.25em}
		\caption{45 \AA~initial pore spacing}~\label{fig:p2p_45}
	\end{subfigure}
	\begin{subfigure}{0.3\textwidth}
		\includegraphics[width=\textwidth]{p2p_50.png}\quad	
		\vspace{-1.25em}
		\caption{50 \AA~initial pore spacing}~\label{fig:p2p_50}
	\end{subfigure}
	\begin{subfigure}{0.3\textwidth}
		\includegraphics[width=\textwidth]{p2p_55.png}
		\vspace{-1.25em}
		\caption{55 \AA~initial pore spacing}~\label{fig:p2p_55}
	\end{subfigure}
	\begin{subfigure}{0.3\textwidth}
	\includegraphics[width=\textwidth]{p2p_diagram.png}
	\caption{}~\label{fig:p2p_diagram}
%		\caption{In a perfect hexagonal, there are 5 distances that should be exactly
%		equal. As pictured in this diagram the distance from 1-2, 1-3, 1-4, 2-4, and 3-4
%		are the same.}~\label{fig:p2p_diagram}
  	\end{subfigure} 
	\caption{The pore spacing during the restrained portion of the
	dry equilibration procedure is shown. Every 50 ps (0.05 ns) we reduce the 
	force constant on the position restraints according to the sequence: 1000000,
	3162, 56, 8, 3, 2, 1, 0 kJ mol$^{-1}$ nm$^{-2}$. When we chose an initial pore
	spacing below experimental value (a) or at the experimental value (b), there
	is an abrupt change in pore spacing when the position restraints are reduced to 56 kJ
	mol$^{-1}$ nm$^{-2}$. When we initially position pores 45 \AA~apart (c) the pore spacing
	remains relatively stable. When we initially place pores 50 \AA~apart (d), the pore
	spacing decreases nearly linearly once the restraints are reduced to 56 kJ
	mol$^{-1}$ nm$^{-2}$. When we initially place pores 55 \AA~apart (e), so that monomers
	do not intersect with adjacent pores, and position restraints are reduced to 56
	kJ mol$^{-1}$ nm$^{-2}$, the pore spacing changes erratically before stabilizing
	when force constants are reduced below 3 kJ mol$^{-1}$ nm$^{-2}$. (f)
	In a perfect hexagonal, there are 5 distances that should be exactly
	equal. As pictured in this diagram the distance from 1-2, 1-3, 1-4, 2-4, and 3-4
	are the same. All plots in (a)--(e) are the average of these distances.} 
	\label{fig:p2p}
  \end{figure}

@@

  %Our model is insensitive to initial configuration if monomer placement is chosen within reason.

  We chose to build all systems using 20 monomers per column. We made this choice
  by considering a number of factors that directly affect the quality of our results. 
  Clearly, choosing a small number of layers will cause the system to suffer from 
  finite size effects. Even in our 20 monomer-per-column system, axial correlations
  between stacked monomers persist throughout the entire correlation function (see
  Figure ~\ref{M-fig:correlation}). The correlation function of the system which we
  built with 40 monomers per column is very similar to that created with 20 monomers
  per columns (continues to oscillate for its full length
  (Figure~\ref{fig:tall_correlation_full}). If we continue adding monomers, the 
  correlation function will eventually fully decay but it is not worth the 
  computational expense. Regardless, the correlation lengths calculated from 20
  monomer per column simulations are in sufficient agreement with experiment.

@@

  \begin{table}[h]
  \centering
  \begin{tabular}{cc}
  \toprule
  System                         & Pore Spacing ($nm$) & Pore Spacing ($nm$)              \\
                                 & Pore centers        & $\dfrac{x-box vector}{2}$ ($nm$) \\ 
  \midrule
  Sandwiched, Ordered            & 4.18 $\pm$ 0.13     & 4.16 $\pm$ 0.01                  \\
  Parallel Displaced, Ordered    & 4.17 $\pm$ 0.12     & 4.18 $\pm$ 0.01                  \\
  Sandwiched, Disordered         & 4.05 $\pm$ 0.14     & 4.01 $\pm$ 0.01                  \\
  Parallel Displaced, Disordered & 4.03 $\pm$ 0.07     & 4.01 $\pm$ 0.01                  \\
  \bottomrule
  \end{tabular}
  \caption{The pore spacing measured in two ways agree within uncertainty. In the main
  text, we report values calculated by measuring the average distance between pore centers. 
  Alternatively, one can estimate the pore spacing as half of the $x$ and $y$ box vectors.
  Both methods agree within uncertainty, but measuring the distance between pore centers
  does a better job of showing the spread of pore-to-pore distances.}~\label{table:p2p}
  \end{table}

  \section{The nematic order parameter}\label{method:nematic_order}

  We calculated the nematic order parameter for our system in order to
  understand the degree of ordering among monomer head groups. Typically, the
  nematic order parameter is calculated for nematic liquid crystal systems which
  are characterized by unidirectional ordering of liquid crystal monomers. The
  preferred direction of monomers is defined by the unit director vector,
  $\mathbf{n}$. Assuming a single preferred direction of alignment, the nematic
  order parameter, $S$, is defined as:
  \begin{equation}
	 S = \frac{1}{2} \langle(3\cos^2\theta -1)\rangle
	\label{eqn:nematic_order_parameter}
  \end{equation}
  where $\theta$ is the angle between the molecular long axis and $\mathbf{n}$.
  In a perfectly ordered nematic liquid crystal system, the molecular axis of each
  monomer should be aligned with $\mathbf{n}$ and give an order parameter of $S=1$. 
  We are interested in quantifying the degree of monomer head group alignment 
  between systems. We use Eq.~\ref{eqn:nematic_order_parameter} to accomplish this
  by defining $\mathbf{n}$ as the z-axis (or pore axis), and then measuring the angles,
  $\theta$, between $\mathbf{n}$ and the vectors perpendicular to the plane of
  the aromatic head groups ($\mathbf{v}$ in Figure~\ref{fig:director}).

  %BJC: change n to v in picture.
  \begin{figure}[!htb]
  \centering
  \includegraphics[width=0.75\linewidth]{nematic_director.png}
  \caption{We calculate the nematic order parameter by measuring the angle
  between $\mathbf{v}$ and the z-axis.}~\label{fig:director}
  \end{figure}

@@

  \begin{figure}[!htb]
  \begin{subfigure}{\linewidth}
        \centering
        \begin{subfigure}{0.45\linewidth}
                \centering
                \includegraphics[width=\linewidth]{layered_nematic_order.png}
                \caption{Sandwiched, Ordered}~\label{fig:sandwich_nematic}
        \end{subfigure}%
        \begin{subfigure}{0.45\linewidth}
                \centering
                \includegraphics[width=\linewidth]{offset_nematic_order.png}
                \caption{Parallel Displaced, Ordered}~\label{fig:offset_nematic}
        \end{subfigure}
        \begin{subfigure}{0.45\linewidth}
                \centering
                \includegraphics[width=\linewidth]{disorder_sandwich_nematic_order.png}
                \caption{Sandwiched, Disordered}~\label{fig:disorder_sandwich_nematic}
        \end{subfigure}%
        \begin{subfigure}{0.45\linewidth}
                \centering
                \includegraphics[width=\linewidth]{disorder_offset_nematic_order.png}
                \caption{Parallel Displaced, Disordered}~\label{fig:disorder_offset_nematic}
        \end{subfigure}
  \end{subfigure}
  \caption{The distribution of angles between nematic director vector (see
	  Figure~\ref{fig:director}) and the z-axis averaged over the equilibrated portion
	  of each trajectory}~\label{fig:nematic_distribution}
  \end{figure}

@@

%  \begin{figure}
%	\centering
%	\begin{subfigure}{0.45\textwidth}
%		\centering
%		\includegraphics[width=\textwidth]{ps5layered.png}
%		\caption{}\label(fig:ps5layered}
%	\end{subfigure}
%	\begin{subfigure}
%		\centering
%		\includegraphics[width=\textwidth]{ps5offset.png}
%		\caption{}\label{fig:ps5offset}
%	\end{subfigure}
%  \end{figure}
  \section{Tail organization}

  In order to understand the origin of R-spots, we looked for ordering among
  the tails. We divided the tail into three sections: tail fronts, tail 
  middles and tail ends. The sections are illustrated in 
  Figure~\ref{fig:tail_sections}.

  \begin{figure}[!htb]
  \centering
  \includegraphics[width=0.5\textwidth]{tail_sections.png}
  \caption{We divided the tail into three regions: tail-fronts (blue), 
  tail-middles (green) and tail-backs (red).}\label{fig:tail_sections}
  \end{figure}

  We plotted the center of masses of each tail section and measured the angle 
  between each center of mass and its nearest neighbor center of masses of the
  same tail section. For example, Figure~\ref{fig:centroids} shows a 3D plot of
  the center of masses of each tail front. We wanted to know average location
  of nearest neighbors to each center of mass. The red bead in 
  Figure~\ref{fig:centroids} is a randomly selected center of mass. The
  green beads are the nearest neighbor center of masses to the red bead. We
  measured the angle between the red bead and each green bead with respect to
  the $xy$ plane. We repeated the same calculation for each center of mass at
  each simulation frame and were able to plot the average distribution shown
  in Figure~\ref{fig:tail_fronts}. We repeated the calculation for the tail
  middles (Figure~\ref{fig:tail_middles}) and the tail ends 
  (Figure~\ref{fig:tail_ends}). The amount of order decreases as we explore
  packing of tails farther from the head group.

@@

@@

>>>>

\externaldocument[M-]{Draft}
\title{Supporting Information : Understanding the nanoscale structure of hexagonal phase lyotropic liquid crystal membranes}
\author{Benjamin J. Coscia \and Douglas L. Gin \and Richard D. Noble \and Joe Yelk \and Matthew Glaser \and Xunda Feng \and Michael R. Shirts}

@@

  We parameterized monomers according to the following procedure:
  \begin{enumerate}

	\item \textit{Create monomer structure file with connectivity} : We
	drew atomistic structures using MarvinSketch
	17.13~\cite{chemaxon_marvinsketch_2017} with all hydrogen atoms drawn out
	explicitly. We optimized the 3D geometry of the structure using the `Clean in
	3D' function of MarvinSketch.  We saved the structure as a .mol file, then
	converted it to .pdb format using Open Babel 2.4.1
	\cite{oboyle_open_2011,noauthor_open_nodate}.

	\item \textit{Assign GAFF atomtypes using \texttt{antechamber}} : Using
	the .pdb structure file as input, we ran
	\texttt{antechamber}~\cite{wang_automatic_2006} using the AM1-BCC charge model.
	The net charge on the monomer is input as -1 since the sodium ion is kept as 
	a separate residue. We use \texttt{LEaP}~\cite{case_ambertools16_2016} and the
	output of \texttt{antechamber} to create Amber topology files. A detailed
	tutorial can be accessed elsewhere \cite{walker_antechamber_nodate}.

	\item \textit{Create GROMACS topologies from Amber output} : The output
	of \texttt{LEaP} is a .inpcrd and a .prmtop file which are Amber topology
	files. Using acpype.py \cite{sousa_da_silva_acpype_2012}, we converted the
	\texttt{LEaP} output into GROMACS .gro and .top files.

	\item \textit{Perform a simulated annealing procedure on the monomer} :
	We created a cubic box around the monomer using the GROMACS command \texttt{gmx
	editconf}. The monomer was centered in the box with edges of the
	box spaced at least 3 nm from the monomer on all sides. We ran an energy minimization
	on the system with the steepest descent algorithm. Next we performed an NVT
	simulated annealing procedure. We linearly decreased the temperature of the
	system from 1000 K to 50 K over the course of 10 ns. We randomly chose a monomer
	configuration from the last 10 \% of the trajectory.

	\item \textit{Reassign charges with \texttt{molcharge}}: With the monomer
	configuration taken from the annealed trajectory, we reassigned charges using
	\texttt{molcharge} with the am1bccsym method in order to ensure charges 
	are symmetric. This condition is not guaranteed with \texttt{antechamber}. 
	The charges in the GROMACS topology file (.top) were replaced with the 
	new charges calculated by \texttt{molcharge}.

	\item \textit{Anneal again to get final structure} : We repeated the
	same simulated annealing procedure using the monomer topology with
	\texttt{molcharge} charges. A random monomer configuration was pulled from the
	last 10 \% of the trajectory and was used to build all assemblies reported
	(Figure~\ref{fig:monomer}).

  \end{enumerate} 
  \clearpage
  \begin{figure}[!htb]
	\centering
        \includegraphics[width=0.9\textwidth]{monomer.png}
	\caption{Atomistic representation of the monomer Na-GA3C11. White atoms
		represent hydrogen, cyan atoms represent carbon, red atoms represent oxygen and
		the blue atom is sodium.}\label{fig:monomer}
  \end{figure}

@@

  We use the nematic order parameter as defined in Section~\ref{method:nematic_order} to 
  observe any progress towards system ordering. In both cases, we see that the nematic
  order parameter stays close to zero for the duration of the simulation time, indicating
  an isotropic arrangement of head groups (Figure~\ref{fig:nematic_isotropic}). The values
  are slightly negative since there are only 400 monomers. 
  Equation~\ref{eqn:nematic_order_parameter} is bounded between -1 and 1, however, with a
  proper choice of nematic director, measured values will range between 0 and 1. With a 
  large number of samples, an isotropic system would have an order parameter of 0.

@@

  We chose to build all systems using 20 monomers per column. We made this choice
  by considering a number of factors that directly affect the quality of our results. 
  Clearly, choosing a small number of layers will cause the system to suffer from 
  finite size effects. Even in our 20 monomer-per-column system, axial correlations
  between stacked monomers persist throughout the entire correlation function (see
  Figure~\ref{M-fig:correlation} of the main text). The correlation function of the system which we
  built with 40 monomers per column is very similar to that created with 20 monomers
  per columns (continues to oscillate for its full length
  (Figure~\ref{fig:tall_correlation_full})). If we continue adding monomers, the 
  correlation function will eventually fully decay but it is not worth the 
  computational expense. Regardless, the correlation lengths calculated from 20
  monomer per column simulations are in sufficient agreement with experiment.

@@

	% BJC: Could put vertical lines to indicate where position restraint values are reduced
  	\begin{figure}[!htb]
	\centering
	\begin{subfigure}{0.3\textwidth}
	\includegraphics[width=\textwidth]{p2p_diagram.png}
	\caption{}~\label{fig:p2p_diagram}
%		\caption{In a perfect hexagonal, there are 5 distances that should be exactly
%		equal. As pictured in this diagram the distance from 1-2, 1-3, 1-4, 2-4, and 3-4
%		are the same.}~\label{fig:p2p_diagram}
  	\end{subfigure} 
	\begin{subfigure}{0.3\textwidth}
		\includegraphics[width=\textwidth]{p2p_39.png}\quad
		\vspace{-1.25em}
		\caption{39 \AA~initial pore spacing}~\label{fig:p2p_39}
	\end{subfigure}
	\begin{subfigure}{0.3\textwidth}
		\includegraphics[width=\textwidth]{p2p_41.png}\quad
		\vspace{-1.25em}
		\caption{41 \AA~initial pore spacing}~\label{fig:p2p_41}
	\end{subfigure}
	\begin{subfigure}{0.3\textwidth}
		\includegraphics[width=\textwidth]{p2p_45.png}
		\vspace{-1.25em}
		\caption{45 \AA~initial pore spacing}~\label{fig:p2p_45}
	\end{subfigure}
	\begin{subfigure}{0.3\textwidth}
		\includegraphics[width=\textwidth]{p2p_50.png}\quad	
		\vspace{-1.25em}
		\caption{50 \AA~initial pore spacing}~\label{fig:p2p_50}
	\end{subfigure}
	\begin{subfigure}{0.3\textwidth}
		\includegraphics[width=\textwidth]{p2p_55.png}
		\vspace{-1.25em}
		\caption{55 \AA~initial pore spacing}~\label{fig:p2p_55}
	\end{subfigure}
	\caption{(a) In a perfect hexagonal, there are 5 distances that should be exactly
	equal. As pictured in this diagram the distance from 1-2, 1-3, 1-4, 2-4, and 3-4
	are the same. All plots in (b)--(f) are the average of these distances and represent
	the pore spacing during the restrained portion of the dry equilibration procedure.
	Every 50 ps (0.05 ns) we reduce the force constant on the position restraints according to the sequence: 1000000,
	3162, 56, 8, 3, 2, 1, 0 kJ mol$^{-1}$ nm$^{-2}$. When we chose an initial pore
	spacing below experimental value (b) or at the experimental value (c), there
	is an abrupt change in pore spacing when the position restraints are reduced to 56 kJ
	mol$^{-1}$ nm$^{-2}$. When we initially position pores 45 \AA~apart (d) the pore spacing
	remains relatively stable. When we initially place pores 50 \AA~apart (e), the pore
	spacing decreases nearly linearly once the restraints are reduced to 56 kJ
	mol$^{-1}$ nm$^{-2}$. When we initially place pores 55 \AA~apart (f), so that monomers
	do not intersect with adjacent pores, and position restraints are reduced to 56
	kJ mol$^{-1}$ nm$^{-2}$, the pore spacing changes erratically before stabilizing
	when force constants are reduced below 3 kJ mol$^{-1}$ nm$^{-2}$. } 
	\label{fig:p2p}
  \end{figure}

@@

  Based on evidence from section~\ref{M-section:rpi} of the main text, we
  believe it is most likely that the head groups stack somewhere between
  the sandwiched and parallel displaced modes which we studied (Figure~\ref{fig:between_pd}).

  \begin{figure}[!htb]
  \centering
  \begin{subfigure}{0.45\textwidth}
  \centering
  \includegraphics[width=\textwidth]{pd_36.eps}
  \caption{}\label{fig:pd_36}
  \end{subfigure}
  \begin{subfigure}{0.45\textwidth}
  \centering
  \includegraphics[width=\textwidth]{pd_lt36.eps}
  \caption{}\label{fig:pd_lt36}
  \end{subfigure}
  \caption{Head groups are likely stacked in a configuration between sandwiched and
  parallel displaced. (a) In our parallel displaced configurations with 5 columns 
  per pore, we placed head groups so the angle, $\theta$, between vectors extending
  from vertically stacked head groups to the pore center equals 36$\degree$ 
  ((360$\degree$ / $ncol$) / 2 , where $ncol$ is the number of columns per pore). 
  (b) In the true configuration, $\theta$ is likely distributed between 0$\degree$
  and 36 $\degree$. When $\theta$=0$\degree$, the system architecture is the same
  as a sandwiched configuration.}\label{fig:between_pd}
  \end{figure}

@@

  \begin{table}[h]
  \centering
  \begin{tabular}{ccc}
  \toprule
  System                         & Pore Spacing ($nm$) & Pore Spacing ($nm$)              \\
                                 & Pore centers        & $\dfrac{x\ box\ vector}{2}$ ($nm$) \\ 
  \midrule
  Sandwiched, Ordered            & 4.18 $\pm$ 0.13     & 4.16 $\pm$ 0.01                  \\
  Parallel Displaced, Ordered    & 4.17 $\pm$ 0.12     & 4.18 $\pm$ 0.01                  \\
  Sandwiched, Disordered         & 4.05 $\pm$ 0.14     & 4.01 $\pm$ 0.01                  \\
  Parallel Displaced, Disordered & 4.03 $\pm$ 0.07     & 4.01 $\pm$ 0.01                  \\
  \bottomrule
  \end{tabular}
  \caption{The pore spacing measured in two ways agree within uncertainty. In the main
  text, we report values calculated by measuring the average distance between pore centers. 
  Alternatively, one can estimate the pore spacing as half of the $x$ and $y$ box vectors.
  Both methods agree within uncertainty, but measuring the distance between pore centers
  does a better job of showing the spread of pore-to-pore distances.}~\label{table:p2p}
  \end{table}

%  \begin{figure}
%	\centering
%	\begin{subfigure}{0.45\textwidth}
%		\centering
%		\includegraphics[width=\textwidth]{ps5layered.png}
%		\caption{}\label(fig:ps5layered}
%	\end{subfigure}
%	\begin{subfigure}
%		\centering
%		\includegraphics[width=\textwidth]{ps5offset.png}
%		\caption{}\label{fig:ps5offset} 
%	\end{subfigure}
%  \end{figure}

  \section{2D Small Angle X-ray Scattering Data}

  \clearpage
  \section{The nematic order parameter}\label{method:nematic_order}

  We calculated the nematic order parameter for our system in order to
  understand the degree of ordering among monomer head groups. Typically, the
  nematic order parameter is calculated for nematic liquid crystal systems which
  are characterized by unidirectional ordering of liquid crystal monomers. The
  preferred direction of monomers is defined by the unit director vector,
  $\mathbf{n}$. Assuming a single preferred direction of alignment, the nematic
  order parameter, $S$, is defined as:~\cite{chaikin_principles_1995}
  \begin{equation}
	 S = \frac{1}{2} \langle(3\cos^2\theta -1)\rangle
	\label{eqn:nematic_order_parameter}
  \end{equation}
  where $\theta$ is the angle between the molecular long axis and $\mathbf{n}$.
  In a perfectly ordered nematic liquid crystal system, the molecular axis of each
  monomer should be aligned with $\mathbf{n}$ and give an order parameter of $S=1$. 
  We are interested in quantifying the degree of monomer head group alignment 
  between systems. We use Equation~\ref{eqn:nematic_order_parameter} to accomplish this
  by defining $\mathbf{n}$ as the $z$-axis (or pore axis), and then measuring the angles,
  $\theta$, between $\mathbf{n}$ and the vectors perpendicular to the plane of
  the aromatic head groups ($\mathbf{v}$ in Figure~\ref{fig:director}).

  %BJC: change n to v in picture.
  \begin{figure}[!htb]
  \centering
  \includegraphics[width=0.75\linewidth]{nematic_director.png}
  \caption{We calculate the nematic order parameter by measuring the angle
  between $\mathbf{v}$ and the $z$-axis.}~\label{fig:director}
  \end{figure}

@@

  \begin{figure}[!htb]
  \begin{subfigure}{\linewidth}
        \centering
        \begin{subfigure}{0.45\linewidth}
                \centering
                \includegraphics[width=\linewidth]{layered_nematic_order.png}
                \caption{Sandwiched, Ordered}~\label{fig:sandwich_nematic}
        \end{subfigure}%
        \begin{subfigure}{0.45\linewidth}
                \centering
                \includegraphics[width=\linewidth]{offset_nematic_order.png}
                \caption{Parallel Displaced, Ordered}~\label{fig:offset_nematic}
        \end{subfigure}
        \begin{subfigure}{0.45\linewidth}
                \centering
                \includegraphics[width=\linewidth]{disorder_sandwich_nematic_order.png}
                \caption{Sandwiched, Disordered}~\label{fig:disorder_sandwich_nematic}
        \end{subfigure}%
        \begin{subfigure}{0.45\linewidth}
                \centering
                \includegraphics[width=\linewidth]{disorder_offset_nematic_order.png}
                \caption{Parallel Displaced, Disordered}~\label{fig:disorder_offset_nematic}
        \end{subfigure}
  \end{subfigure}
  \caption{The distribution of angles between nematic director vector (see
	  Figure~\ref{fig:director}) and the $z$-axis averaged over the equilibrated portion
	  of each trajectory. The ordered sandwiched configuration has a higher nematic order
	  parameter than the ordered parallel displaced configuration. The disordered
	  sandwiched configuration has a higher nematic order parameter than the 
	  disordered parallel displaced configuration.}~\label{fig:nematic_distribution}
  \end{figure}

@@

  \section{The effect of imposed tail tilt}\label{section:tail_tilt}

  Monomer tails prefer to lie flat rather than tilt with respect to the
  membrane plane. We created a monomer whose tails tilted ca. 30$\degree$. 
  We added position restraints to the ends of the monomer tails with the
  same force constant as the head groups (see section~\ref{section:equilibration}).
  As we slowly reduce the force constants, the tails begin to tend towards
  a negligible tilt angle (Figure~\ref{fig:tail_tilt}).

  \begin{figure}[!htb]
  \centering
  \includegraphics[width=0.5\textwidth]{tilt_cat.png}
  \caption{As we reduce the force constants applied to the end of each
  tail of an initially tilted monomer configuration, the average tilt 
  angle with respect to the membrane plane decreases until it is nearly
  negligible.}\label{fig:tail_tilt}
  \end{figure}

  \section{Tail organization}\label{section:tail_packing}

  In order to understand the origin of R-spots, we looked for ordering among
  the tails. We divided the tail into three sections: tail-fronts, tail-middles
  and tail-ends. The sections are illustrated in Figure~\ref{fig:tail_sections}.

  \begin{figure}[!htb]
  \centering
  \includegraphics[width=0.5\textwidth]{tail_sections.png}
  \caption{We divided the tail into three regions: tail-fronts (blue), 
  tail-middles (green) and tail-ends (red).}\label{fig:tail_sections}
  \end{figure}

  We plotted the center of masses of each tail section and measured the angle 
  between each center of mass and its nearest neighbor center of masses of the
  same tail section. For example, Figure~\ref{fig:centroids} shows a 3D plot of
  the center of masses of each tail front. We wanted to know the average location
  of nearest neighbors to each center of mass. The red bead in 
  Figure~\ref{fig:centroids} is a randomly selected center of mass. The
  green beads are the nearest neighbor center of masses to the red bead. We
  measured the angle between the red bead and each green bead with respect to
  the $xy$ plane. We repeated the same calculation for each center of mass at
  each simulation frame and were able to plot the average distribution shown
  in Figure~\ref{fig:tail_fronts}. We repeated the calculation for the tail
  middles (Figure~\ref{fig:tail_middles}) and the tail ends 
  (Figure~\ref{fig:tail_ends}). The amount of order decreases as we explore
  packing of tails farther from the head group.

@@

  \section{Carboxylate Dihedrals}

@@

  \section{Autocorrelation of monomer dihedrals}

  In section~\ref{M-section:slow_dynamics} of the main text, we calculated the
  autocorrelation function of the outer tail either dihedrals pictured
  in Figure~\ref{fig:dihedrals}.
